---
title: "Introduction to the Used-Habitat Calibration Plots (uhcplots) Package"
author: "Althea ArchMiller and John Fieberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to uhcplots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


This package contains functions that allow the user to: 

-  recreate the simulation examples in Fieberg, J.R., Forester, J.D., Street, G.M., Johnson, D.H., ArchMiller, A.A. and Matthiopoulos, J., 2018. Used‐habitat calibration plots: a new procedure for validating species distribution, resource selection, and step‐selection models. Ecography, 41(5), pp.737-752.
 
- create Used-Habitat Calibration (UHC) plots via cross-validation

We begin by providing a short overview of the steps involved in creating UHC plots.  We then illustrate the functions in the *uhcplots* package with a simple simulation example from Fieberg et al. (in review).  Specifically, we show how UHC plots can be used to explore the impact of a missing predictor using data splitting.

Other vignettes illustrate how UHC plots can be used to:

- explore the impact of a missing predictor using spatially-stratified cross-validation
- validate step-selection functions using a model fit to moose location data from Fieberg et al. (in review)


## How to calculate UHC plots

UHC plots are simple graphical tools that can be used to validate species distribution, 
habitat selection, and step-selection models. UCH plots compare distributions
of habitat covariates at the used (or presence) points to distributions predicted
based on a model fit to independent (training) data.  

To create a UHC plot, we:

1. Split the data into training and test data sets.  This may be accomplished by
data splitting - either randomly or using temporal or spatial stratification (e.g.,
the model may be fit to a subset of sites or years, with other 
sites or years used as "out-of-sample" data).  Alternatively, one can use bootstrapping
or cross-validation to evaluate out-of-sample predictions.  We provide an example
using spatially-stratified cross-validation later on in this vignette.

2. Summarize the distribution of habitat covariates at the used (i.e., presence) points in the test data set, $f^u(z)$. In our examples, we use a kernel density estimator to represent $f^u(z)$.  Similarly, summarize the distribution of the habitat covariates at the available (i.e., background)  points in the test data set, $f^a(z)$.  Differences between these two densities signal that the covariate will be an important predictor of the species distribution. 

3. Fit a model to the training data set.  Store $\hat{\beta}$ and $\widehat{\mathrm{cov}}(\hat{\beta})$ to characterize the uncertainty in the parameters (ignoring the intercept if using logistic regression). Assuming we have a large enough sample for $\hat{\beta}$ to be approximately normally distributed, we can draw samples from a multivariate normal distribution, $\mathrm{MVN}(\hat{\beta}, \widehat{\mathrm{cov}}(\hat{\beta})$), to account for uncertainty in the estimated parameters.  This uncertainty may alternatively be captured using a non-parametric bootstrap or via samples from a posterior distribution (if implementing the model in a Bayesian framework). We will refer to the distribution capturing uncertainty in $\hat{\beta}$ as the *joint parameter distribution* to recognize that this will be a multivariate distribution if more than one covariate is included in the model.

4. Do the following $M$ times (with loop index $i$):

    a. To account for parameter uncertainty, select new vector of parameter values randomly from their joint parameter distribution, $\beta^i$.
    b. Estimate the relative probability of selection for the test data: $w(x^{test}\beta^i)= \exp(x^{test}\beta^{i})$. 
    c. Select a simple random sample of $n^{test}_u$ observations from the combined (presence and background) test data, with probabilities of selection proportional to $w(x^{test}\beta^i)$ from step [3b]; here $n^{test}_u$ is the number of used points in the test data set. 
    d. Summarize the distribution of habitat covariates associated with the points chosen in step [3c], $\hat{f}^u(z)_i$ . 

5. Compare the observed distribution of covariate values at the presence points, $f^u(z)$ from step [1], to the predicted distribution of these characteristics, $\hat{f}^u(z)_i$  from step [3], across the $M$ simulations. We illustrate two plotting options:  

    a. Option 1:  overlay $f^u(z)$ (from step [1]) on a 95\% simulation envelope constructed using the $\hat{f}^u(z)_i$. 
    b. Option 2: plot the 2.5$^{th}$ and 97.5$^{th}$ quantiles of $f^u(z)-\hat{f}^u(z)_i$.    


## UHC plots to explore the impact of a missing predictor using data splitting

This simulation example comes from Fieberg et al. (in review). In this example, we will assume that the probability of selection is 
proportional to $\exp(0.5x_1 - x_2)$, where $x_1$ is elevation and $x_2$ is 
precipitation. We begin by loading 2 packages:
 
- KernSmooth (for creating smooth density estimates of habitat covariates)
- uchplots (for creating UHC plots!)

```{r libraries, warning=F,message=F}
library(devtools)
library(KernSmooth)
library(uhcplots)
set.seed(1000)
```


We will simulate used and available pionts using the *uhcdatasimulator* function in the *uhcplots* package.  This function was created to allow users to rerun the examples from Fieberg et al. (in review), but with the possibility of modifying:

- numbers of used and available locations in training and test data sets
- the magnitude of the regression coefficients determining the relative probability of selection as a function of elevation and precipitation (example = "missing predictor") or as a function of temperature and temperature$^2$ ( example = "non-linear")
- the correlation among elevation and precipitation in the test and training data sets (if running example = "missing predictor")

Below, we replicate the example illustrating UCH plots in the presence of a missing predictor (precipitation) for the scenario where the correlation between elevation and precipitation was = 0.3 in the training data and -0.3 in the test data.

Example-specific settings:

```{r example1, warning=F,message=F}
# Number of "used" locations
nused_train <- nused_test <- 100 
# Number of "available" or background locations
navail.train <- navail.test <- 10000
# Number of simulated available locations
ntemp <- 1000000
# Number of simulations used to create UHC plot (i.e., "M" from manuscript)
nsims <- 1000
# Example name
example <- "missing predictor"
# Example betas
betas <- c(0.5, -1)
# Labels for x-axes
xlabs <- c("Elevation","Precipitation")
```

Scenario-specific settings:
```{r scenario1}
# correlation (elevation, precipitation) in training and test data
corx.train <- 0.3                    
corx.test <- -0.3                    
# Plot labels (for the correlations, above)
labcor1 <- expression(rho[list(x[1],x[2])]==0.3)
labcor2 <- expression(rho[list(x[1],x[2])]==-0.3) 
textplot <- ""                    
# panel lables
panlabs1 <- c("A)","B)")     
panlabs2 <- c("A)","B)", "C)", "D)")
# Add a title to the top if top <-1
top <- 1                           
```

**Simulate Data**

Simulate training data using the uhcdatasimulator() function.
```{r sim1atrain}
traindat <- uhcdatasimulator(nused = nused_train,
                             navail = navail.train,
                             betas = betas,
                             corx = corx.train,
                             ntemp = ntemp,
                             example = example)
```

Simulate test data using the uhcdatasimulator() function.
```{r sim1atest}
testdat <- uhcdatasimulator(nused = nused_test,
                             navail = navail.test,
                             betas = betas,
                             corx = corx.test,
                             ntemp = ntemp,
                             example = example)
```

**Fit GLM Models to Training Data**

Now, we fit two models, one with both $x_1$ and $x_2$ (i.e., "correct" model) 
and one with just $x_1$ (i.e., "missing predictor" model).

```{r glm1a}
# Correct model
train.correct <- glm(y ~ elev + precip,
                     family = binomial,
                     data = traindat)
# Missing predictor model
train.missingprecip <- glm(y ~ elev,
                           family = binomial,
                           data = traindat)
```

**Create Predicted Distributions Using the *uhcsim* Function**

Next, we create simulation envelopes for the environmental characteristics (elevation and precipitation) 
at the observed locations in the test data using both models. As in the paper, we create 1000 predicted distributions (i.e., we set M = 1000 in step 4 of the process). Note, the *uhcim* function requires the following arguments:

- nsims = number of predicted distributions to create (this is analogous to "M" in step 4)
- nused_test = number of used points to select in the test data set
- xmat = design matrix for the fitted species distribution model
- fit_rsf = fitted logistic regression model 
- z = matrix of habitat covariates to cross-validate.

This function returns a (nsims, $n^{u}_{test}$, number of covariates = ncol(z)) array containing the habitat covariates associated with each of $n^{u}_{test}$ randomly selected locations  across all M simulations.  
```{r uhcsim1}
# Correct model
xhat.correct <- uhcsim(nsims = 1000, 
                       nused_test = nused_test,
                       xmat = testdat[, c("elev","precip")], 
                       fit_rsf = train.correct, 
                       z = testdat[, c("elev","precip")])


# Missing predictor model
xhat.missingprecip <- uhcsim(nsims = nsims, 
                             nused_test = nused_test, 
                             xmat = as.matrix(testdat[, c("elev")]), 
                             fit_rsf = train.missingprecip, 
                             z = testdat[, c("elev","precip")])
```

**Summarize the Predicted Distributions Using the *uhcdenscalc* Function**

Next, we calculate a kernel density estimate of the predicted distribution of both predictors for each of the M data sets, $\hat{f}^u(x_1)_i$ and $\hat{f}^u(x_2)_i$ ($i = 1, ..., M$). The *uhcdenscalc* function requires the following arguements:

- rand_sims = an (nsims, $n^{u}_{test}$, number of covariates) array containing the habitat covariates associated with each of $n^{u}_{test}$ randomly selected locations  across M separate simulations.  Usually, this array will be generated by the *uhcsim* or *uhcsimstrat* function.
- dat = the characteristics associated with observed locations in the test data set
- avail = the characteristics associated with available locations in the test data set

The *uhcdenscalc* function will return a list of density estimates associated with:

- all M sets of predicted locations (object\$densrand);
- the observed locations (object\$densdat); and
- the available locations (object\$densavail)

```{r uchdenscalc}
# Correct model
# uhcdenscalc function for both predictors
denshats.elev.correct <- uhcdenscalc(rand_sims = xhat.correct[,,1],
                                  dat = subset(testdat, y==1, select="elev"), 
                                  avail = subset(testdat, y==0, select="elev")) 
denshats.precip.correct <- uhcdenscalc(rand_sims = xhat.correct[,,2], 
                                     dat = subset(testdat, y==1, select="precip"),
                                     avail = subset(testdat, 
                                                  y==0, select="precip")) 

# Missing predictor model
# uhcdenscalc function for both predictors
denshats.elev.missingprecip <- uhcdenscalc(rand_sims = xhat.missingprecip[,,1],
                                           dat = subset(testdat, 
                                                      y==1, select="elev"), 
                                           avail = subset(testdat, 
                                                        y==0, select="elev")) 
denshats.precip.missingprecip <- uhcdenscalc(rand_sims = xhat.missingprecip[,,2], 
                                             dat = subset(testdat, 
                                                        y==1, select="precip"), 
                                             avail = subset(testdat, 
                                                        y==0, select="precip"))
```

### Create UHC plot (formatted for manuscript)

Lastly, we use the *uhcdensplot* function to produce the UHC plot (as illustrated in Fieberg et al. in review). This function requires as input:

- densdat = the kernel density estimates of observed points in the test data set
- densrand = the kernel density estimates for the habitat covariate at the predicted test data points (across M predicted data sets)
- includeAvail = an indicator determining whether the distribution of the available locations should be drawn on the plot
-  densavail = the kernel density estimates of available points in the test data set (needed if includeAvail = TRUE)


```{r plotuhc00, fig.width=7}
par(mfrow=c(1,5), mar=c(2,2,2,2), oma=c(3, 0, 5, 0), bty="L")

# Use the first plot for text
    plot(1, type="n", xlab="", ylab="", 
         xlim=c(0, 1), ylim=c(0, 1), 
         xaxt='n',yaxt='n',bty='n')
    text(0.01, 0.9, "Training Data", cex=1, adj=0)
    text(0.2, 0.72, labcor1 , cex=1,adj=0) 
    text(0.01, 0.52, "Test Data", cex=1, adj=0)
    text(0.2, 0.40, labcor2 , cex=1, adj=0) 
    legend(0.01, 0.35, c("Available", "Used", "Predicted"), 
            lty = c(1, 2, 1), col = c("red", "black", "grey"), 
            bty = "n", lwd = c(1, 1, 5))

# Incorrect model (missing predictor)
uhcdensplot(densdat = denshats.elev.missingprecip$densdat, 
                densrand = denshats.elev.missingprecip$densrand,
                xl = c(-4,9),
                includeAvail = TRUE,
                densavail = denshats.elev.missingprecip$densavail,
            includeLegend = F) 
    mtext(outer=F, side=2, line=3, "Density", cex=1)
    mtext(side=3, line=1,  panlabs2[1], cex=1, ad=0)
    mtext(outer=F, side=1, line=4, xlabs[1], cex=1.0)
    
    uhcdensplot(densdat = denshats.precip.missingprecip$densdat, 
            densrand = denshats.precip.missingprecip$densrand,
            xl=c(-9,6),
            includeAvail = TRUE, 
            densavail = denshats.precip.missingprecip$densavail,
            includeLegend = F) 
    mtext(outer=F, side=1, line=4, xlabs[2], cex=1.0)
    mtext(side=3, line=1, panlabs2[2], cex=1, ad=0)

# Correct model (both predictors)
uhcdensplot(densdat = denshats.elev.correct$densdat, 
            densrand = denshats.elev.correct$densrand,
            xl = c(-5,9),
            includeAvail = TRUE, 
            densavail = denshats.elev.correct$densavail,
            includeLegend = F)  
    mtext(side=3, line=1, panlabs2[3], cex=1, ad=0)
    mtext(outer=F, side=1, line=4, xlabs[1], cex=1.0)

uhcdensplot(densdat = denshats.precip.correct$densdat, 
            densrand = denshats.precip.correct$densrand,
            xl = c(-9,3.5),
            includeAvail = TRUE, 
            densavail = denshats.precip.correct$densavail,
            includeLegend = F) 
    mtext(side=3, line=1, panlabs2[4], cex=1, ad=0)
    mtext(outer=F, side=1, line=4, xlabs[2], cex=1.0)

    if(top==1){
      mtext(outer=T, side=3.5, line=2, adj=0.80, 
            expression(y %~% elev+precip), cex=1)
      mtext(outer=T, side=3.5, line=2, adj=0.40, 
            expression(y %~% elev), cex=1)
      }
```


### Alternative UHC Plot

We also illustrate an alternative plot, depicting the 2.5$^{th}$ and 97.5$^{th}$ quantiles of $f^u(z)-\hat{f}^u(z)_i$, using the *uhcdiffdensplot* function 

```{r uhcdiff00, fig.width=7}
par(mfrow=c(1,5), mar=c(2,2,2,2), oma=c(3, 0, 5, 0), bty="L")

# Use the first plot for text
plot(1, type="n", xlab="", ylab="", 
     xlim=c(0, 1), ylim=c(0, 1), 
     xaxt='n',yaxt='n',bty='n')
text(0.01, 0.9, "Training Data", cex=1, adj=0)
text(0.2, 0.72, labcor1 , cex=1,adj=0) 
text(0.01, 0.52, "Test Data", cex=1, adj=0)
text(0.2, 0.40, labcor2 , cex=1, adj=0) 

# Incorrect model (missing predictor)
uhcdiffdensplot(densdat = denshats.elev.missingprecip$densdat, 
                densrand = denshats.elev.missingprecip$densrand,
                xl = c(-5,9))  
    mtext(outer=F, side=2, line=3, "Density", cex=1)
    mtext(side=3, line=1,  panlabs2[1], cex=1, ad=0)
    mtext(outer=F, side=1, line=4, xlabs[1], cex=1)

uhcdiffdensplot(densdat = denshats.precip.missingprecip$densdat, 
                    densrand = denshats.precip.missingprecip$densrand,
                    xl=c(-9,3.5)) 
    mtext(outer=F, side=1, line=4, xlabs[2], cex=1)
    mtext(side=3, line=1, panlabs2[2], cex=1, ad=0)

# Correct model (both predictors)
uhcdiffdensplot(densdat = denshats.elev.correct$densdat, 
                    densrand = denshats.elev.correct$densrand,
                    xl = c(-4,9))  
    mtext(side=3, line=1,  panlabs2[3], cex=1, ad=0)
    mtext(outer=F, side=1, line=4, xlabs[1], cex=1)

uhcdiffdensplot(densdat = denshats.precip.correct$densdat, 
                    densrand = denshats.precip.correct$densrand,
                    xl = c(-9,6))
    mtext(side=3, line=1,  panlabs2[4], cex=1, ad=0)
    mtext(outer=F, side=1, line=4, xlabs[2], cex=1)

if(top==1){
  mtext(outer=T, side=3.5, line=2, adj=0.80, 
        expression(y %~% elev+precip), cex=1)
  mtext(outer=T, side=3.5, line=2, adj=0.40, 
        expression(y %~% elev), cex=1)
}
```
