---
title: "Used-Habitat Calibration Plot Examples"
author: "Althea ArchMiller and John Fieberg"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction 

# UHC Approach

# Examples

## Example 1: Missing predictors

In this example, we will assume that the probability of selection is 
proportional to $\exp(0.5x_1 - x_2)$, where $x_1$ is elevation and $x_2$ is 
precipitation. For all scenarios, data were generated under the following
scenario: $\text{var}(x_1) = \text{var}(x_2) = 4$ and the correlation between
$x_1$ and $x_2$, or $\rho(x_1,x_2)$, varies for the training and test data sets. 

Example-specific settings
```{r example1}
library(devtools)
load_all()
set.seed(1000)
# Number of "used" locations
nused_train <- nused_test <- 100 
# Number of "available" or background locations
navail.train <- navail.test <- 10000
# Number of simulated available locations
ntemp <- 1000000
# Number of simulations used to create UHC plot (i.e., "M" from manuscript)
nsims <- 1000
# Example name
example <- "missing predictor"
# Example betas
betas <- c(0.5, -1)
# Labels for x-axes
xlabs <- c("Elevation","Precipitation")
```

### 1a. Scenario where $\rho(x_1,x_2)$ = 0 for both training and test data sets.

Scenario-specific Settings
```{r scenario1}
corx.train <- 0                      
corx.test <- 0                       
scenario <- "corr00"                 
labcor1 <- expression(rho[list(x[1],x[2])]==0)
labcor2 <- expression(rho[list(x[1],x[2])]==0) 
textplot <- ""                       
panlabs1 <- c("A)","B)")     
panlabs2 <- c("A)","B)", "C)", "D)")
top <- 1                           
outfig1 <- "uhcplots/graphics/Boyce1a.pdf"
outfig2 <- "uhcplots/graphics/crosssim1a.pdf"
```

**Simulate Data**

Simulate training data using the uhcdatasimulator() function.
```{r sim1atrain}
traindat <- uhcdatasimulator(nused = nused_train,
                             navail = navail.train,
                             betas = betas,
                             corx = corx.train,
                             ntemp = ntemp,
                             example = example)
```

Simulate test data using the uhcdatasimulator() function.
```{r sim1atest}
testdat <- uhcdatasimulator(nused = nused_test,
                             navail = navail.test,
                             betas = betas,
                             corx = corx.test,
                             ntemp = ntemp,
                             example = example)
```

**Fit GLM Models**

Now, we fit two models, one with both $x_1$ and $x_2$ (i.e., "correct" model) 
and one with just $x_1$ (i.e., "missing predictor" model).

```{r glm1a}
# Correct model
train.correct <- glm(y ~ elev + precip,
                     family = binomial,
                     data = traindat)
# Missing predictor model
train.missingprecip <- glm(y ~ elev,
                           family = binomial,
                           data = traindat)
```



Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
